\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{a4wide}
\usepackage{url}
\usepackage[auth-sc]{authblk}

\title{\sc Deska: Tool for Central Administration of a Grid Site}
\author{Kundrát, J.}
\author{Kerpl, L.}
\author{Krejčová, M.}
\author{Hubík, T.}
\affil{Institute of Physics of the AS CR, Prague, Czech Republic}

\date{October 1, 2010}

\begin{document}

\maketitle
 
\begin{abstract}
Running a Tier-2 WLCG site is a demanding task, especially given the ever
increasing number of machines, auxiliary hardware and accompanying software
packages.  System administrators have always tried to find ways to reduce their
efforts while not compromising systems' security and reliability.  In this
paper, we focus on reducing the duplication of information describing a complete
datacenter, including the relations between various components.  We provide
details about the Deska project, an in-house CLI appliance for managing these
metadata.
\end{abstract}


\section{Introduction}

Recent developments in the hardware area are bringing unmatched challenges to
datacenter operators.  The ever increasing computational power density, along
with advancements in semiconductor engineering and mass production, enables
integrating many hundreds of CPU cores into a single rack while maintaining
reasonable TCOs.  Machines which formerly occupied several rack units are now
produced in tiny blade form factors.  That all leads to a rapid rise of the
number of services which have to be managed and taken care of.

This rapid rise is, however, usually unmatched by a corresponding growth in the
size of the IT department.  A concrete example about the situation in Prague is
presented in a site report given by Tomas Kouba \cite{kouba}. Therefore, system
administrators seek ways to mitigate many of the common problems they are
facing.  One way of possible problem mitigation
scenario is making sure that all information one might need is conveniently
accessible, kept up to date and actually used in production.

Design of such a system has to fulfil certain requirements, from being easy to
use to scaling well with growing number of nodes.  Most of the existing hardware
inventory systems were deemed unusable for production use at Prague, for reasons
provided later in the article.  Therefore, a new tool called Deska was
designed, and its architecture and design considerations are provided in this
document.  The goal was to provide a tool which would integrate with the existing
infrastructure as smoothly as possible, yet be universal enough to keep up with
future demands.  We have to merge best practices from traditional relational
database design with data versioning, conflict resolution and changeset merging.

When the Deska project is mature and tried in production, we have to focus on
possible ways to further integrate it with day-to-day operation.  As an example,
a reasonable way of putting the data to good use is enabling a one-action
deployment or re-installation of new machines or using the Deska's knowledge of
physical features of the machines to visualize the server room and provide power
usage estimates.


\section{The way to automation}

Since its establishment in 2002, the Prague Tier-2 WLCG site has grown from 32
single-core machines into the 2010's biggest site in the Central Europe,
offering their users more than 2600 modern CPU cores.  This capacity, hosted by
the Institute of Physics of the AS CR, is nowadays dedicated to performing
heavy-duty HPC computations, supporting both in-house users from particle and
solid-state physics, as well as to taking part in international collaboration
for experiments in FNAL, USA and CERN.

Such a rapid rise in computing capacity was not achieved just by hardware
advancements, but is also a result of deploying more and more machines. In fact,
the Prague data center occupation was risen several times, from the
original 32 machines to its current status of 336 computing nodes and tens of
auxiliary or service nodes.

This increase in the computing capacity was not, however, compensated by a
matching grow in staffing operators.  While the original size was handled by
three full-time employees, the current situation is actually a bit worse.  The
total contracted capacity is slightly more than three full-time-equivalents
(FTE), but only two people are dedicated to day-to-day operations, and the rest
consists of three students, each with 0.25 to 0.5 FTE employments.  Some
assistance is provided by other section from the Institute, but it should be
obvious that the amount of work and responsibility that each of the employees
has to bear has risen considerably over the years.

A naive manager might consider such situation a pleasant one -- the people
should not be slacking while at work, after all.  However, due to the nature of
the IT industry, it is in fact desired not to operate the staffers at full
utilization levels all the time.  When they do not have time for reading
newspapers and are kept busy with day-to-day operation, these duties will
inevitably suffer when an emergency arises.  Therefore, the administrators
simply {\em have} to have free time in order to handle unexpected situations.

In their quest to make their life easier, people at the Institute have deployed
many tools which were supposed to automate the day-to-day operation.  At first,
in-house scripts were used for performing changes to the nodes, typically in the
form of a shell {\tt for} loop.  This approach did not scale at all, and
required manual supervision for completion.  It also caused trouble when some
nodes were offline at the time of the change, perhaps because of a hardware
trouble.  When such host got powered up again, the change wasn't performed on
it, and there was nothing to remind the responsible people about that.

Therefore, yet more in-house scripts were written to look after the above
mentioned scripts' execution.  The scalability problems were not mitigated at
all, and there were many corner cases not really fully understood, nor properly
addressed.  Therefore, a way to migrate to an ``intelligent solution'' was
clearly needed.

We evaluated numerous tools, from Cfengine version 2 and Puppet to specialized
applications developed inside the HEP community like Quattor.  Ultimately, we
decided that Cfengine would fit our goals in the best way, and therefore chose
it.  A proper tool for automation should ideally deal with random host changes
and just check the final state to see if it matches the specified contract.  For
example, a correct way of converting a machine to use an LDAP database for user
accounts, as done by Cfengine, is the following:

{\scriptsize
\begin{verbatim}
{ /etc/nsswitch.conf
    BeginGroupIfNoLineMatching ^passwd:([[:space:]])+files([[:space:]])+ldap$"
      HashCommentLinesStarting "passwd:"
      Append "passwd: files ldap"
    EndGroup
}
\end{verbatim}
}

When compared to a more-usual {\tt sed} magic, it is obviously more verbatim,
but also arguably much more readable.  Using Cfengine properly certainly takes
some time and effort, but the results should pay off pretty soon -- the
configuration can be stored in a single place which could easily be put under a
version control system (VCS) like Subversion or Git, and Cfengine's design
ensures that a change, or rather a description of a desired {\em state}
specified once will be verified for eternity, and if it can't be reached, an
alarm would be risen.

Using Cfengine along with LDAP for centralized management of user accounts and
Kickstart/Anaconda for new hosts deployment cut a big part of the burden from the
sysadmins' shoulders, but still left much to be dealt with.  The most annoying
problem, as perceived by the Institute's staffers, is the need to specify the
same thing on several places.

\section{Information duplication}

In order to better illustrate what we mean with duplication of information,
let's consider a typical use case of new machine deployment.  After a
procurement is finished, the technicians from the winning bidder arrive
on site along with the actual hardware.  The machines and supporting technology
get physically installed, cables are laid out and the electricity is switched
on.  Staffers have to write down certain information, from rack locations to
part numbers and warranty information.  After that is done, the network switches
have to be properly configured, the MAC addresses previously gathered have to be
entered into the DHCP server's database, IP addresses and matching hostnames
have to be assigned and registered in the DNS system etc.  When the machines are
turned on for the first time, a proper Kickstart file has to be delivered over
the PXE environment, which has to be set up properly.  When the operating system
is installed, a configuration management system, Cfengine in our case, has to be
told which role the machine is going to fulfill, so that it can configure
everything accordingly.  Last but not least, various monitoring appliances have
to be made aware of the new machine's existence, so we get to know about
possible problems before we suffer availability or reliability loss, or our
users notice.

Therefore, the information about a host's existence is needlessly duplicated
into several places.  In the case of the Prague's Tier-2 site, some of these
places are the following:

\begin{itemize}
    \item HW inventory DB
    \item Warranty \& issue tracking
    \item Switch port configuration
    \item DHCP server
    \item DNS
    \item Cfengine roles
    \item Torque's CPU multipliers
    \item MRTG \& RRD network graphs
    \item Nagios
    \item Ganglia
    \item Munin
    \item \ldots
\end{itemize}

It is clear that this duplication will only lead to troubles over time.  It
isn't entirely obvious that an innocent operation like warranty replacement of a
motherboard shall result in changing HW inventory database, warranty database,
switch configuration, the DHCP server, and also network intrusion detection
system.  What is needed here is a {\em central place} to store all host
metainformation.

When such a place is ready, we have to put it to real use, too.  Patching
existing services to query this place, or a database, is likely not an option,
so tools have to be written for pushing this information to places like network
switches or DHCP servers' configuration.  A slight delay has to be anticipated,
but in reality it does not complicate things much.  Distribution of these
configuration files is outside of scope of the Deska project, and will usually
be handled by standard tools like Puppet or Cfengine.

\section{Requirements}

Based on their experience with another in-house attempt at a central inventory
database, a web-based tool, the staffers decided to assign the biggest
importance to the ease of use of the new tool.  As all of the staffers are also
heavy Linux users, they tend to prefer the command-line interface to more
traditional web-based ones.

Numerous situations have shown that having a textual dump of a database is of
much experience.  There are simply times when everything is down, perhaps as a
result of fatal electricity incident or on-location fire.  When the network is
offline, having a full copy of the database, perhaps without much history, but
at least with all the data, is a hard requirement.  If such a dump resides on
each administrator's laptop's hard drive, chances are that people will be able
to consult up-to-date information when performing critical services recovery,
reducing the possibility of making fatal mistakes and causing more damage by
accident.

The Cfengine experience illustrated, among other things, the usefulness of
keeping track of history of changes.  When it is possible to identify what
changes got performed by whom and what was their reason, people tend to make
less mistakes, and, what is most important, they are less likely to repeat them.

Finally, while most web-based tools would work in an ``online'' mode where all
changes get pushed to the production repository immediately, or perhaps after
pressing the ``go live'' button, a two-phase commit with manual review of
changes is much more common in the realm of version control systems like
Subversion or Git.  Under such system, changes are permanently copied only after
you have seen them one more time, usually as a {\tt diff}, a difference report
between the former and the new version.  In the case of the inventory database,
it is worthwhile to display changes in the database along changes in various
services' configuration.

\section{Deska design and operation}

The design of the Deska's UI was inspired by the way datacenter Ethernet
switches are configured, most notably by the Cisco IOS shell.  Therefore, the
user is presented with a command line interface sporting intelligent
tab-completion, and they are supposed to navigate through a hierarchy of
objects, similar to what she would do when configuring a port of the switch.

These objects model a real-world entities which are found in a typical machine
room, as well as abstract entities representing concepts like a ``brand'' or a
``model'' of a particular machine.  A general rule of thumb is that before one
can instantiate a ``physical object'', a matching ``type'' has to be defined --
for example, before one adds fifty Altix XE320 WNs, the {\tt xe320} hardware
model has to be defined.

To date, the following {\em abstract} entities are supported:

\begin{description}
    \item[boxmodel] A type of a physical box, like a room, universal rack, blade
        enclosure or a twin server chassis
    \item[hwmodel] A type of a HW box, from generic types like a 1U server to
        well-specified entities like ``HP DL360''
\end{description}

Matching the above, the corresponding {\em instances} of a specified kind are:

\begin{description}
    \item[rack] An instance of a {\tt boxmodel}.  Contrary to what the name
        might suggest, this statement could represent anything from a machine
        room to physical rack, one concrete instance of a blade enclosure or
        even one of many twin server chassis which are deployed in a real-world
        rack.
    \item[host] An instance of a physical computer.  This computer is always a
        physical device, and runs exactly one operating system image.  Support
        for virtualized machines is available, but not discussed in this
        document.
\end{description}

In order to better illustrate the above concepts, and to provide a real-life
example about how they are used, consider the following snippet of the
configuration of the Prague site:

{\scriptsize
\begin{verbatim}
# At first, we define a type representing plain room in a building
boxmodel room
    outer width 1000000
    outer height 1000000
    outer depth 1000000
end

# Next, we define a template for a regular rack:
boxmodel generic-rack
    # ...it is supposed to be placed in a room defined above
    in room

    # Because the outer element, ie. ``room'' in this case, does not define any bays,
    # we specify physical dimensions:
    outer width 700
    outer depth 1300
    outer height 2000

    # Contrary to ordinary rooms, a rack is usually divided into areas with well-defined boundaries.
    # Let's call them ``bays'' here and specify how they are organized:
    bay height 44 order bottom-up start-at 1
    # ...which is syntactically equivalent to:
    # bay
    #    height 44
    #    order bottom-up
    #    start-at 1
    # end
    # ...and after adding default values, to this:
    # bay
    #    height 44
    #    width 1
    #    depth 1
    #    order bottom-up
    #    start-at 1
    # end
end
\end{verbatim}
}

We have seen that there are many syntactic ways to express the same concept; we
hope that such an approach is not confusing.

In real world, certain models are a {\em specialization} of other items.  We
have already defined a generic rack with some default dimensions.  Suppose a new
delivery comes in racks with slightly different dimensions:

{\scriptsize
\begin{verbatim}
# A slightly bigger rack...
boxmodel APC-rack
    # ...is a specialization of the generic-rack...
    template generic-rack

    # ...slightly taller...
    outer height 2300

    # ...and all other properties, including the bay organization, are inherited from
    # its parent, the ``generic-rack''
    bay height 48
end
\end{verbatim}
}

This specialization is called a {\tt template}.  A template works simply by
inheriting all properties from the parent item, with the possibility to override
each and every one of them.

The reason for differentiating between a {\em type} of an object and {\em
properties} of one particular instance is to force some consistency on how the
operators work with the database.  If it was allowed to change arbitrary
properties on the {\em instance} level, people would likely abuse that
possibility to avoid using constructs like the {\tt boxmodel} altogether.

Now that we have defined our abstract building blocks, let's see how to
instantiate real objects which are found in the machine room:

{\scriptsize
\begin{verbatim}
# Start with the ``room'', which is not derived from anything
rack machine-room
    model room
end

# Add real racks now
rack L01
    model APC-rack
    in machine-room # a reference to a ``real object'' defined above

    # ``machine-room'' is of instance ``room'' which does not provide any bays,
    # therefore we have to fall back to absolute positioning:
    position x 10 y 20 z 0
end

# As a further example of enclosure nesting, suppose this blade chasses:
rack hp-enc-p-1
    model hp-blade-p-chassis # ...this would have to be defined somewhere

    # This instance is a blade enclosure which fits to a rack. We also know its
    # dimension in rack units (ie. we know the number of occupied ``bays in a
    # rack'' at this point), we only have to place it somewhere. The placement is
    # done via a ``significant corner'', which is the bay number of the lowest,
    # left-most and front-most bay occupied by the box.
    in L01 bay 10
end
\end{verbatim}
}

We have already demonstrated how to define a {\tt boxmodel} and a {\tt rack}.
Matching counterparts for the physical machines are {\tt hwmodel} and a {\tt
host}.  We'll start by showing how to use the {\tt hwmodel}, along with yet
another template demonstration, and follow by a {\tt host} definition:

{\scriptsize
\begin{verbatim}
hwmodel 1u
    in generic-rack
    # no bay occupation indicated -> default to "1 bay"
end

hwmodel dl360g5
    template 1u
end

hwmodel dl140
    template 1u
    benchmark hepspec 8
    cpu ``Intel Xeon 3.06 GHz''
    sockets 2
    cores-per-socket 1
    logical-cores-per-physical-core 2
    disk ``ATA 80GB``
    power 110W
end

host ha1
    hw dl360g5
    serial 1234567890
    rack L01 bay 0
end
\end{verbatim}
}

As we can see, even though certain vital information, like the network layout,
is still missing from the table, we already have all the data needed for drawing
a ``map'' of where machines are located.

In order for the database to be more valuable and to be able to be used as a
source of data for various services, it is crucial to introduce network
definitions, too.  It is done in the following way:

{\scriptsize
\begin{verbatim}
network wn-nat
    vlan 172
    ip 172.16.0.0/16
end
\end{verbatim}
}

The purpose of maintaining these data is to be able to perform certain checks on
data validity, like verifying that a host's IP address belongs to the assigned
range.  Now let's introduce a first real server:

{\scriptsize
\begin{verbatim}
host hypericum10
    # Housekeeping records
    # Serial No. of the whole box
    serial J019MF6C2J
    # Warranty No. & expiration date
    warranty KE1421631007 expires 2006-12-12
    purchased 2004-05-03

    # Physical dimension, architecture, location and interconnects
    hw dl140
    rack L04 bay 40
    kvm unit 4 port 7
    interface eth0 mac 00:0f:20:7a:e7:9c net wn-nat ip 172.16.4.10 switch swL041 port 5

    # Logical roles
    role wn
end
\end{verbatim}
}

This concept of lists of object of well-known types can be extended to include
other entities commonly found in a datacenter, should the need arise.

\section{Comparison to existing tools}

To the best of our knowledge, the Deska project is the only tool which focuses on
a CLI interface.  There are indeed numerous other applications, from the
``industrial standard'' OCS Inventory to less known projects like the
RackMonkey, as well as appliances developed inside the WLCG community, like
Quattor fabric management tool or the Smurf database.

In addition to a missing CLI, the mainstream way of tackling the inventory
management problem is usually focused on automatic discovery of existing
properties.  This is a very useful approach when working with desktop computers,
which often change outside of the network manager's approval, but could be less
optimal when working in a well-controlled datacenter environment.  In addition,
while it could be desirable to automatically pick certain changes in the
environment and propagate them to the database, at least under certain
conditions, the Deska tool is meant to provide an {\em authoritative source} of
information -- what is in the database is the desired situation, representing
the sysadmins' decisions, and should be implemented in the real world and not
vice versa.

\section{Future work}

The design described so far presented the database scheme which was pretty much
set in stone.  We recognize that each site's requirements are different, and
have ourselves struggled with providing a good approximation of the real
situation in Prague and translating it to the database structure.  An
intermediate solution for a request for the possibility of customization is to
allow the administrator to specify certain additional attributes to attach to
existing types of entities in the database.  However, such an approach would
make expressing constraints rather difficult, at least due to the extension
fields' unspecified structure.

The considerations above led us to another approach within the Deska
application.  In future versions, we are going to make the DB structure
modifiable by the administrator.  We will retain the basic idea that the
database consists of lists of objects of a well-defined type, and that each such
object has a list of attributes, or properties, of one of a few predefined
types.  Such objects are entities like the {\tt rack} or a {\tt boxmodel}
defined above.  The definitions of such objects, along with a list of supported
attributes and their data types, should be easily provided by the administrator.
The Deska project will of course ship with a predefined set of objects which are
common in the datacenter.

In consistency with the Deska's motto of reducing duplication of information, we
are seeking ways to define the DB scheme at exactly one place, along with all
possible constraints.  We do not have any motivation to invent yet another
language for describing the database, but plan to use SQL DDL as the way to
define the DB structure, along with some rules for coding conventions.  Our goal
is not to create an ``über Database'' capable of describing each and every
situation, but to come up with reasonable limits allowing the administrators to
describe their infrastructure accurately.

\section{Conclusion}

We have described some of the current griefs which the staffers at the Institute
of Physics of the AS CR are going through, along with a proposed solution which
could reduce this burden.  We have learned valuable lessons when trying to
design a proper database structure, using the Prague Tier-2 WLCG site as an
example.  The database structure which we picked is centered on lists of objects
of a few predefined types.

During the implementation works, we have discovered that it makes a lot of sense
to allow the administrators to extend the database structure.  Therefore, most
of the hard-coded limitations of the original Deska design were lifted.  The
Deska will now read the database structure at the startup and attempt to
reconstruct the valid object types and their attributes.  In practise, such an
approach to the problem helps to eliminate a lot of boiler-plate code.  What
needs to be done, however, is defining a reasonable set of constraints for the
administrators to follow.


\subsection{Acknowledgement}

The work on the Deska project was partially funded by the Institute of Physics
of the~AS~CR,~v.v.i.  The author of this poster can be reached at {\tt
kundratj@fzu.cz}.

\begin{thebibliography}{9}
    \addcontentsline{toc}{chapter}{References}
    \bibitem{kouba}Kouba, T. et al: The Prague WLCG Tier-2 Site Report. November
        2010.
\end{thebibliography}

\end{document}

