% vim: spelllang=en spell textwidth=120
\documentclass[deska]{subfiles}

\begin{document}

%FIXME: change to appendix
\chapter{Existing Tools}
\label{sec:evaluating-existing-tools}

Many proprietary tools were excluded from the evaluation due to their high price or inability to effectively work in a
vendor-neutral manner.\footnote{It is common for the manufacturers of server-class hardware to have their own solution
to mass management of the individual nodes.  However, having full support for working with devices from other equipment
manufacturers is not usual.}  We have nonetheless evaluated a number of open-source tools, both from the general IT
industry, as well as those which were developed at other WLCG sites.

FIXME: references

\section{Industry-standard Tools}

\subsection{OCS Inventory NG}

The OCS Inventory NG is perhaps the most well-known information system used in the context of hardware management.  The
biggest selling point of the OCS Inventory is its ability to {\em enumerate} the whole network and automatically
discover all of the connected devices, and with the help of locally-installed agents monitor various aspects of each
host's operation.  On the other hand, it does not appear to be designed for the Institute's use case, which is {\em
providing} an authoritative answer based on policy decisions instead of an transient snapshot of a datacenter state.

\subsection{Rackmonkey}

The Rackmonkey is an opensource project whose main goal is to make hardware management easy.  Unfortunately, at the time
we were evaluating the existing tools, the Rackmonkey had no support for machines with a different form factor than
``pizza boxes'', like blades or the twin servers, and no features which could be used for managing other kinds of
equipment.

\subsection{IBM xCat}

IBM xCat, the ``Extreme Cluster Administration Tool'',\footnote{The current meaning of the abbreviation is ``Extreme
{\em Cloud} Administration Tool, apparently to better match current buzzwords in the industry.} is a promising project
with a philosophy which is very close to the traditional, old-school Unix system administrator.  At the same time, it is
optimized for central {\em deployment} and {\em management} of the individual nodes, which are tasks that are already
well handled at the Institute.


\section{WLCG-specific Alternatives}

\subsection{Quattor}

Quattor is an opensource project with roots in the WLCG grid community.  The biggest advantage of this tool is its
diverse community which nowadays include developers from other industries, including several financial institutions.
However, the project has rather wide scope and --- similarly to the xCat --- delivers a complete, integrated solution
for multiple tasks, including fabric management.

\subsection{Smurf}

The Smurf project is another tool which is produced by the WLCG community.  Developed in the CC-IN2P3 in Lyon, France,
it sports a modular architecture and appears to be well thought-through.  Unfortunately, at the time we were evaluating
the existing rojects, the Smurf project has been undergoing a major overhaul and there were essentially no
documentation.

\section{Augmenting What Works Well: Nagios, Ganglia, Cfengine, Puppet,\ldots}

There are many existing opensource tools which assist in the day to day operation of a massive computing infrastructure;
some of them are intended as monitoring appliances, others aim to streamline the propagation of changes from a central
place to individual nodes.  Deska does not plan to replace these tools, but to {\em leverage} their power and use them,
instead of reinventing the wheel.


\end{document}
