% vim: spelllang=en spell textwidth=120
\documentclass[deska]{subfiles}
\begin{document}

\chapter{Objects and Relations}

\begin{abstract}
The Deska database can store data of varying structure.  In this chapter, we provide an overview of what can be stored
and how to use the provided facilities to design a database scheme which closely matches the real-world processes.  The
basic building blocks of the Deska database, the {\em kinds}, {\em objects} and {\em relations}, are described here.
\end{abstract}

\section{Objects and Relations: The Basic Building Blocks}
\label{sec:objects-and-relations}

Deska does not try to deny its background in relation databases, and therefore it comes at no surprise that the basic
building blocks, the Deska {\em objects}, {\em kinds} and {\em relations}, all have their well-known parallels in the
world of relational databases and object-oriented programming.

Anything which is stored in the Deska database is an {\em object}. An object can be thought of as a row in a DB table ---
all objects of the same type, or {\em kind} as we call that in the Deska jargon,\footnote{Unfortunately we were unable
to call Deska kinds {\em classes}, which would be the best name for them, because a lot of languages use {\tt class} as
a keyword.  We have chosen to use the slightly unintuitive word {\tt kind} instead of obvious alternatives like {\tt
class\_}.} live in one table, and they all can
store the same amount of information in their {\em attributes}, the database columns.  One of the columns is usually a
primary key, which in the Deska corresponds to a special attribute called {\em name}, and serves as a unique identifier
of an object.

Two objects from different kinds can have the same name, and unless the scheme defines a special connection between
these two kinds, the two objects will remain completely isolated from each other.  There is no inherent hierarchy
between the Deska kinds, all of them exist at the same level.  Even though the real-world is full of examples where
something ``belongs'' to something else, like an {\tt eth0} interface could belong to machine called {\tt srv06},
there's no such thing at this level of the Deska interface --- when you want to work with a particular object, perhaps
the {\tt srv06}'s {\tt eth0}, you always have to identify it through a globally unique name.  Of course, all of this is
hidden (and much else provided) by the upper layers thanks to the concept of {\em relations}.

\section{Relations}
\label{sec:objects-and-relations-relations}

A {\em relation}~\footnote{These relations do {\em not} refer to relations from relational algebra, the name is again a
result of a limited set of words usable as identifiers.  We have deemed the word ``relation'' to be more accurate than
the alternatives like ``association'' or ``connection'', which all have well-defined meaning in certain programming
frameworks.} is an extra association between two individual object instances which can be used to model wildly differing
connections between real-world objects.  A relation is defined by the left and right {\em kinds} to which it refers and
by the associated {\em type} of the relation.\footnote{This is further augmented by auxiliary information of which
object attribute is used to determine the target object of the relation, but all of that can be safely ignored for now.
Don't worry, we'll get back to that later.}

The relations serve different purposes, all of which are described in this section.  Some of them are suitable for
maintaining referential integrity, others serve as ``context helpers'' for humans to work with, yet others can be used
to build complex inheritance and composition patterns, common to the object-oriented programming.

\subsection{Referential Integrity}

Let's start with perhaps one of the simplest relation, the {\tt REFERS\_TO} one.  The sole purpose of this relation is
to make sure that an object's attribute does not contain just an ordinary text, but that the value of said attribute
refers to an object of a given kind.

There are many use cases for such relation, the most basic one is probably the DB-way of defining enumerations.  Suppose
we are designing a hardware inventory and want to keep track of vendors or manufacturers.  It would be impractical to
modify the database code when a new company selling servers emerges, so it makes sense to treat the list of
manufacturers as data.  This is simply done by creating a new kind, let's call it ``vendor'', and instantiating new
objects for each supplier currently in use.  The {\tt machine} kind which typically holds objects representing each
physical server would then gain a new attribute called {\tt vendor} with type {\tt identifier}, which will make sure
that the possible values have to conform to the syntax constraints defined for identifiers.  For the referential
integrity to work, a new {\em relation} shall be created.  This new relation shall have the type {\tt REFERS\_TO}, and
be from the kind {\tt machine} to kind {\tt vendor}.

When a new delivery from a brand new company arrives, activating that new manufacturer is simply a matter of
creating new object of the {\tt vendor} kind.

\subsection{Relations Explained}

Each relation therefore merely specifies that objects of a particular kind {\em can} somehow refer to objects of some
other kind (or of the same kind, for that matter).  It is important to understand that this relation is {\bf optional},
and can be present or absent on a per-object basis, and that these relations concern the individual objects.  Under the
hood, relations are implemented as database foreign keys, and as foreign keys can be {\tt NULL} unless explicitly
prohibited, an object can easily exist without that particular relation taking any effect.

The foreign key is just a table column similar to other object attributes.  It has its own name defined by the scheme
architect, and is usually fully accessible to the DBAPI users.  The established Deska conventions call for either using
a name which matches the purpose of the relation (like {\tt template}, {\tt inherit} etc.), or simply reusing the name
of the target kind.  There are certain situations (which are even present in the sample schemes, as shipped with Deska)
which warrant using a completely different name, so implementors shall refer to their best judgement when inventing
naming conventions.

\subsection{Multi-value References}

So far, we have shown relations which generally represent an N:1 mapping --- an object of the source kind (a row in the
source table) refers either to nothing, or to exactly one target object.  However, this concept is not sufficient for a
particularly compelling feature, notably for {\em tagging}.

Tagging is probably a well-known concept from the general-purpose computing experience. An object can be tagged by a set
of labels, which generally refer to identifiers of other objects.  In real life, a photo gets tagged with names of
people who are on the picture.  In a data center, each server gets tagged with several roles that it fulfils; for
example, machine {\tt srv012} can act as a web server and an MX host at the same time, and hence shall get two tags, the
{\tt www} and {\tt smtp} ones.

In Deska, this feature is provided by the combination of the {\tt identifier\_set} attribute data type and the {\tt
REFERS\_TO} relation.  In our example, the {\tt role} is a standalone kind, and the {\tt host} kind has an attribute
{\tt services} of type {\tt identifier\_set}.  Where the RDBMS equivalent of tagging would involve dealing with an
auxiliary table just for the mapping, the Deska DBAPI provides an abstraction without any extra kinds.

\subsection{Templates: Default Value Inheritance}

There are certain cases where many objects have certain shared properties.  As an example, imagine a situation where you
order a delivery of 50 new, identical servers.  A lot of their properties is going to be the same, or almost the same,
among all of them.  Therefore it might make sense to create one artificial object, let's call it a {\em template} one,
put all the shared data in there, and have fifty new individual objects ``inherit'' the default values from it.  This
scenario is assisted by the {\tt TEMPLATIZED} relation.

Templates are typically implemented through two kinds, let's call them {\tt hardware} and {\tt template\_hardware}, and
a relation of type {\tt TEMPLATIZED}, which would typically be ``activated'' via the {\tt template} attribute of the
first kind (remember, relations always exist between individual objects).  So in our sample scenario when a vendor
delivers fifty new RM-666 boxes, we would create one instance of the {\tt template\_hardware} kind with the name {\tt
rm666\_2011\_03}, assign interesting properties to that object (like the end of warranty period, support contract number
etc.), and then create fifty {\tt hardware} objects for the individual machines, setting their {\tt template} attribute
to {\tt rm666\_2011\_03}.  This will make sure that attributes which have not been explicitly set at the level of each
object are inherited from the associated template.

This default value substitution can be overridden at any time by setting an explicit non-null value at the level of the
leaf object.  Resetting the attribute value back to {\tt null} later effectively restores the inherited default.  It is
not possible to override an inherited attribute value by {\tt null}; doing that requires explicitly cancelling the
relation for that particular object (in our case, the inheritance is stopped by setting the {\tt hardware}'s attribute
{\tt template} to {\tt null}).

Users should be careful to not abuse the template relation where composition would be more appropriate.  As a rule of
thumb, if it makes sense to deal with a subset of object's attributes as a different entity (maybe distinguishing
between the hardware's physical dimensions, its inner compounds like CPU configuration, the model number, and machine's
rack location), it is usually recommended to go with composition or something else instead of templating.

\subsection{Composition}

In the previous section, we have briefly mentioned that it makes sense to ``split'' properties of one real-world object
into several groups.  On the other hand, doing so imposes a significant burden on the hands of a system administrator
who suddenly does not have to create just a single object for each machine in the rack, but several of them --- one for
the hardware inventory, another for ``rack placement'' (or position), yet another for the logical computer that runs on
top of that, and so on.  To confine this object explosion and to limit the amount of work required to handle such
objects, Deska provides the {\tt CONTAINS} relation.

Using this relation, a logical object is essentially composed of many smaller objects, where each of the ``contained''
parts provides a particular function.  For example, a logical ``computer'' object could be composed of the following
building blocks: {\tt host} (for stuff like hostname and embedded network interfaces), {\tt hardware} (for stuff like
serial number) and {\tt placement} (for the physical space occupied in a rack).  An Ethernet switch would typically
include the {\tt placement} object as well, so it means that the low-level building blocks can be freely reused,
provided no conflicts arise.  It is important to realize that this reusing happens at the {\em kind} level and not at
the object level ---  an individual {\tt placement} object can be linked to a {\tt computer} or to a {\tt switch}, but
never to both.  The {\tt placement} {\em kind}, however, contains many individual objects, some of which refer to {\tt
switch}, and others referring to {\tt computer} instances.

The Deska database contains special hooks which prevent possible ambiguities. For example, when user creates an object
which can be included in another one (in our example, a {\tt placement}), appropriate foreign keys will immediately tie
the freshly created instance to the pre-existing object with the same name, if one exists.  Therefore, if there is
already a switch called {\tt foobar} and one attempts to create a {\tt placement} of the same name, it will immediately
get tied to said switch.  Constraint checks are in place that forbid creating a {\tt switch} and a
{\tt host} with the same name, as it would be impossible to {\em place} them into a rack any time later.  This check is
transitive, i.e. if there is a {\tt machine} which includes {\tt hardware} which in turn includes a {\tt placement}, it
is forbidden to create a {\tt machine} and a {\tt switch} with the same name.

It is up to the Deska administrator to define further database constraints to prevent creating the ``low-level''
building blocks objects on their own right, without being ``used'' by the other objects --- an example where this would
be very appropriate is our {\tt placement} kind.  In other situations however, it might be wise to allow the low-level
objects to exist without actually being referenced from elsewhere (it should generally be allowed to put a machine into
inventory even though it remains off and without an assigned hostname, and hence just the {\tt hardware} object shall
exist, without the accompanying {\tt host}).  In all cases, though, the Deska server automatically takes care of setting
up foreign keys properly when the ``contained'' and ``using'' objects come into existence at any time, no matter which
one gets created first.

When an object at either side of the relation gets renamed, the relation is automatically torn apart and the same rules
which are active at the time of object creation are checked again, enforcing that objects with the same name (and only
those) are linked together.  This means that in order to rename a ``logical'' object which is composed of several other
objects, one has to perform a \deskaFuncRef{renameObject} on all of them.  If a single object gets renamed, the rename
can potentially cancel the relation and immediately establish a link to another instance due to them having the same new
name.

Due to technical reasons, the {\tt CONTAINS} relation is complemented by a {\tt CONTAINABLE} relation in the opposite
direction.  As a rule of thumb, if there is a relation {\em A contains B}, there is also a {\em B containable A}.

\subsection{Object Embedding}

Earlier in this chapter, we have mentioned that there is no structure among the kinds, and that as a result, all objects
have to bear a name which is unique among all objects in the given kind.  However, not only the real world in all its
complexities, but also a simple data center is full of examples where this naming convention would get in the way; the
above mentioned example of a network interface name is a prime example.  There are many ways out of that; the most
obvious ones involve GUIDs and other forms of human-unfriendly means.  In Deska, we have chosen another approach which
is assisted by the {\tt EMBED\_INTO} relation, and is hopefully more readable to humans.

An embedded object ``almost'' violates the limitation that an object's name has to be unique in a given kind.  Instead,
the global name is composed of two parts: the name of the {\em parent} object and a logical name, which has to be valid
only in the context of a parent, which are then joined together by a separator {\tt ->}, hereby forming a unique
identifier.  In practice, {\tt srv012}'s network interface {\tt eth0} would be therefore called {\tt srv012->eth0}.
Under the hood, this is implemented by two distinct attributes and the standard rules for constraint checking guarantee
that there is no potential for possible inconsistency.

In practice, this relaxed naming is still not enough to be directly usable by people, and hence the Deska CLI (and any
other GUIs shipped with Deska, for that matter) fully supports object embedding using this relation.  There is no need
to manually type composed identifiers, as the object nesting is presented transparently to users.

The object embedding also has a feature that allows generating locally unique names for the local part of the name.
That features comes handy when a DB scheme uses object embedding e.g. for tracking failures and other events associated
with a piece of hardware.  The Deska way of handling this use case is to create a kind (let's call it {\tt failure}) and
have its objects embedded into the parent object instances (like {\tt hardware}).  However, doing so without an
additional assistance provided by the CLI or the Deska server would impose an annoying burden on users, who would have
to come up with a unique name when logging a ``failure'' of an equipment.  To address this limitation, the Deska DBAPI
call for creating new objects (\deskaFuncRef{createObject}) has an option for automatically assigning a locally-unique
name for the object just created.  This means that the user just has to ``create a failure for machine {\tt srv012}'',
and the Deska DB itself will assign an identifier like {\tt srv012->1} to it.  This feature is directly available from
the CLI, and the user does not have to deal with low-level naming conventions at all.

\subsection{Relations Rationale}

We have seen that the relations allow describing complex object hierarchies without violating the basic Deska design
rule that the database contains flat lists of objects, which themselves consist of a key-value pairs holding the actual
data.  Applications are therefore free to use the Deska DBAPI without having any knowledge about the relations at all,
and they would still have access to all the data.  At the same time, applications which want to go further and provide
additional assistance can use advanced features (and more function calls) to effectively streamline the end users' work
flow.  The worst thing that the former, ``naive'' applications could encounter are cases where a particular action
(like, for example, creating a {\tt switch} and a {\tt host} with the same name) is denied by the integrity checking
constraints.  On the other hand, applications which choose to employ the relations can use them to build rich object
hierarchies and presenting them to the users without creating much confusion.

\subsection{Limits of Relations}

There are certain rules that limit the possible relations.  Basically, any setup which wishes to use relations has to
respect the following two simple conditions:

\begin{itemize}
    \item {\em No duplicates.} Given kinds A and B and an attribute C of kind A, there is at most one relation between A
        and B that uses attribute C.
    \item {\em No conflicts.} Any setup which might introduce ambiguity to the possible interpretation of relations is
        prohibited.
\end{itemize}

While the first condition is clear on its own, the second rule deserves further explanation.

First of all, there could ever only be a single {\tt EMBED\_INTO} relation between two given kinds, and that relation
must not be present in both directions.  In addition to that, each kind can only be embedded into at most one other
kind, but never more.\footnote{This limitation might be lifted in future versions of Deska.  Applications are advised to
be ready for such a change.}  Finally, this hierarchy must not ever form a circular graph (with kinds being vertices and
relations being edges).

The rules for the {\tt REFERS\_TO} are much more relaxed --- essentially, only the {\em no duplicates} rule applies here.
This relation does not convey any semantic which is important for the Deska tools, the meaning is completely up to the
user and can be paraphrased as just ``this object makes a reference of some type to that object''.

In contrast to the {\tt REFERS\_TO}, the {\tt TEMPLATE} relation is much stricter.  Each kind can only have at most one
kind as the target of the templating relation (ie. the source of the default values).  Additionally, a kind which serves
as a template to some other kind is automatically a template source for itself, and cannot be used as a template by any
other kind.

Semantics of the {\tt CONTAINS} relation are also rather strict.  A low-level building block can be contained in many
higher-level objects, provided that the containment graph never contains a diamond-like shape.  This means that if kinds
A and B both contain kind C, either directly or indirectly, through other kinds, it is forbidden to have another kind D
contain A and B at the same time, again directly or indirectly.

\section{Attributes}

In the previous sections, we have often used the term {\em attribute} without properly codifying what we refer to,
relying on the intuitive definition.  For most purposes, such a definition is enough, but let's reiterate what an
attribute really is and what data it can hold.

As we have already mentioned, a {\em kind} specifies which attributes can be set (and hold data) for an object.  These
attributes are simply a pair of a key and value.  Each attribute hence has a {\em name}, which is defined at the kind
level, and a {\em value}, which is specific to each object instance.  The value is well-typed, meaning that each kind
defines a data type for each attribute, and the data being stored in the attribute values has to conform to the
specified data type.  It is also possible to place additional constraints on the attribute values, either through the
object relations, or via further database triggers.

The list of supported data types covers the most obvious and common needs (like strings and numeric values), and also
includes datacenter-specific types like IPv4/IPv6 or Ethernet MAC addresses.  For a full list of the defined data types,
please consult \secref{sec:json-data-types-reference}.

\end{document}
